Differential invariant signatures for planar Lie group transformations of images

### 2019-09-11 paper/newpaper.tex
minor

The method of moving frames enables one to identify invariants under the action of groups of transformations. It has a long history, as summarised in \cite{Olver14}; particularly noteworthy contributions are by Cartan~\cite{Cartan37} and Olver and co-authors; see~\cite{Olver14} for an overview, and note particularly for an application to~\cite{Calabi98}. The method is described in detail in~\cite{OlverCIT}r.

@@

Olver uses the group $SE(2)$ on curves as an example~\cite{Olver}, recovering the Euclidean curvature as a second-order invariant. We demonstrate the method with two groups: the M\"obius group ($PSL(2,\mathbb{C})$) and the projective group ($PSL(3,\mathbb{R})$).

>>>>

The method of moving frames enables one to identify invariants under the action of groups of transformations. It has a long history, as summarised in \cite{Olver14}; particularly noteworthy contributions are by Cartan~\cite{Cartan37} and Olver and co-authors; see~\cite{Olver14} for an overview, and note particularly for an application to~\cite{Calabi98}. The method is described in detail in~\cite{OlverCIT}.

@@

Olver uses the group $SE(2)$ on curves as an example~\cite{Olver}, recovering the Euclidean curvature as a second-order invariant. As an example, we use it to compute invariants to $SE(2)$ on the plane, before moving on to two other planar Lie group examples: the M\"obius group ($PSL(2,\mathbb{C})$) and the projective group ($PSL(3,\mathbb{R})$).

\subsection{$SE(2)$}

Ignoring translation, the action of $SE(2)$ is simply rotation by some angle $\theta$, i.e.,

\begin{equation}
\left( \begin{array}f_x \\ f_y \end{array} \right) = \left( \begin{array} \cos \theta & \sin \theta \\ - \sin \theta & \cos \theta \end{array} \right) \left( \begin{array} \bar f_{\bar x} \\ \bar f_{\bar y} \end{array} \right).
\end{equation}

Using $c$ for $\cos \theta$ and $s$ for $\sin \theta$ and prolonging the group to second order via the chain rule gives:

\begin{equation}
\arraycolsep=1.4pt
\left(
\begin{array}{c}
  \bar f_x \\ \bar f_y \\ \bar f_{xx} \\ \bar f_{xy} \\ \bar f_{yy} 
 \end{array}
 \right)
 = 
\left(
\begin{array}{ccccc}
 c & -s & 0 & 0 & 0 \\
 s & c & 0 & 0 & 0 \\
0 & 0 & c^2 & -2cs & s^2 \\
0 & 0 & cs & c^2 - s^2 & -cs \\
0 & 0 & s^2 & 2cs & c^2 \\
\end{array}
\right)\left(
\begin{array}{c}
f_x \\ f_y \\ f_{xx} \\ f_{xy} \\ f_{yy} 
 \end{array}
 \right).
\label{eqn:SE2}
 \end{equation}

As $SE(2)$ does not include scaling, lengths stay the same, and so an appropriate choice for first-order normalisation is $\bar f_x = \sqrt(f_x^2 + f_y^2),\:\: \bar f_y = 0$, which sets $c = f_x / \| \nabla f \|$, $s = - f_y / \| \nabla f \|$. This provides the invariant $I_1 = \nabla f$. As there are no more free parameters, the expressions for the second-order terms can be read off as:

\begin{align}
I_2 &= f_x^2 f_{xx} + 2 f_x f_y f_{xy} + f_y^2 f_{yy} \\
I_3 &= f_x f_y (f_{yy} - f_{xx}) + f_{xy} (f_x^2 - f_y^2) \\
I_4 &= f_x^2 f_{xx} - 2 f_x f_y f_{xy} + f_y^2 f_{yy} \\
\end{align}

However, summing $I_2$ and $I_4$ we can write the expression as $f_x^2 f_{xx} + f_y^2 f_{yy}$. Which is the product of 2 invariants in Einstein summation, but a pain in the arse any other way. But I can't see how else to get $f_{ii}$!
